### 2023-06-22 python_excel_ite.tex
working on intro

Despite numerous calls for and predictions of the eradication of spreadsheets from the business world, Excel and other spreadsheets are alive and well. Yes, spreadsheets have well known limitations. Spreadsheets are not databases and probably should not be used as one, especially for relational data models. There are better data visualization and exploration tools. There are better dashboarding tools that are more easily maintained in a production environment. Spreadsheet errors are a huge problem. Version control and documentation present challenges. There are data size limitations and external data links present their own set of challenges. But, spreadsheets are still very well suited to ``what if?'' type modeling due to their flexibility, numerous built in functions, and their integrated sensitivity analysis, goal seeking and visualization capabilities. And of course, they are ubiquitous and familiar to hordes of analysts and modelers.

@@

Numerous R and Python based data science courses have been developed and offered at universities and online learning platforms such as Coursera, EdX, Udemy and others. Some of these courses have made a large amount of their course content freely available for self-learners and for other educators. At our institution, we created a course called Practical Computing for Data Analytics (PCDA) in 2015 that introduced both R and Python (along with a little Linux) for use in data analytics work. The course is taught using combination of R Markdown documents \cite{bibid} and Jupyter notebooks \cite{bibid}. All of the course content is freely available from a publicly accessible course web site \url{URL}.

The PCDA course focuses mostly on data wrangling, exploratory data analysis (EDA), and statistical/machine learning based predictive models. It does include a Python based simulation of the Monte Hall 3-door problem. In teaching EDA using tools like ggplot2, I would challenge students to think about how they would go about trying to do some of the plots we were making in Excel. There is also some limited coverage of using Python to open Excel files and do some sheet manipulation. However, there was no real attempt to do the kind of modeling that was the focus of spreadsheet based modeling textbooks.

Python has continued to make inroads to analytical territory that was formerly dominated by Excel. The widely used Python data analysis library, Pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{bibid}. However, Python was mainly used for data wrangling, visualization and predictive modeling using various flavors of regression and tree based models. The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed greatly to making Python more accessible to business analytics practitioners through high quality posts that were extremely practical and useful. But, it got me to thinking. What about those things for which Excel is well suited such as building formula based models and doing sensitivity analysis on these models? What would those look like in Python?

>>>>

Despite numerous calls for and predictions of the eradication of spreadsheets from the business world, Excel and other spreadsheets are alive and well. Yes, spreadsheets have well known limitations. Spreadsheets are not databases and probably should not be used as one, especially for relational data models. There are better data visualization and exploration tools. There are better dashboarding tools that are more easily maintained in a production environment. Spreadsheet errors are a huge problem. Version control and documentation present challenges. There are data size limitations and external data links present their own set of challenges. But, spreadsheets are still very well suited to ``what if?'' type modeling due to their flexibility, numerous built in functions, and their integrated sensitivity analysis, goal seeking and visualization capabilities. Excel is widely used for building and using models of business problems
to explore the impact of various model inputs on key outputs. Built in
tools such as \href{https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b}{Data
	Tables} and
\href{https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7}{Goal
	Seek} are well known to power spreadsheet modelers.

@@

Numerous R and Python based data science courses have been developed and offered at universities and online learning platforms such as Coursera, EdX, Udemy and others. Some of these courses have made a large amount of their course content freely available for self-learners and for other educators. At our institution, we created a course called Practical Computing for Data Analytics (PCDA) in 2015 that introduced both R and Python (along with a little Linux) for use in data analytics work. The course is taught using a combination of R Markdown documents \cite{bibid} and Jupyter notebooks \cite{bibid}. Jupyter notebooksLiterate computing \cite{Jupyter_proposal}. Educational narratives that support iterative exploration and development. The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact. Jupyter used by many universities as a key component of their educational technology stack.All of the course content is freely available from a publicly accessible course web site \url{URL}.

The PCDA course focuses mostly on data wrangling, exploratory data analysis (EDA), and statistical/machine learning based predictive models. It does include a Python based simulation of the Monte Hall 3-door problem. In teaching EDA using tools like ggplot2, students are challenged to think about how they would go about trying to use Excel to create some of the plots that are easy to do with R and Python based plotting libraries. There is also some limited coverage of using Python to open Excel files and do some sheet manipulation. However, there was no real attempt to do the kind of modeling that was the focus of spreadsheet based modeling textbooks.

Python has continued to make inroads to analytical territory that was formerly dominated by Excel. The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{bibid}.

Python is widely used for data wrangling, visualization and predictive modeling using various flavors of regression and tree based models. The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed greatly to making Python more accessible to business analytics practitioners through high quality posts that were extremely practical and useful. But, it got me to thinking. What about those things for which Excel is well suited such as building formula based models and doing sensitivity analysis on these models? What would those look like in Python?
